---
title: Niche expansion via acquired metabolism facilitates competitive dominance in
  planktonic communities
author: "Veronica Hsu, F. Pfab, and Holly V. Moeller"
date: "12/15/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Main Text: Figure 2
We analyzed the affect of light (at 0, 50, 100, 200 μmol quanta m-2 s-1) and treatment (monoculture vs competition) on max populations and growth rates of Colpidium and Paramecium bursaria. We performed a Tukey's HSD test to determine statistically significant differences at p < 0.05, indicated by different letters. 

The analysis below was used to generate Main Text Figure 2. 

Preparation: Load the data
```{r}
library(readr)
library(readr)
library(forcats)
library(ggplot2)
library(Rmisc)
library(dplyr)
library(tidyr)
df <- read_csv("AcqMet_Tukey.csv")

df$`Light_Level`<- as.character(df$`Light_Level`)
df$Treatment <- as.character(df$Treatment)
df$`Cells/mL` <- as.double(df$`Cells/mL`)
df$Growth_Rate <- as.double(df$Growth_Rate)

Pbur <- df[df$Species == "P bur", ]
Colp <- df[df$Species == "Colp", ]
```

##Tukey's HSD for growth rate of P. bursaria 
```{r}

# Compute the analysis of variance
res.aov_Colp_GR <- aov(`Growth_Rate` ~ `Light_Level`*Treatment, data = Colp)
res.aov_Pbur_GR<- aov(`Growth_Rate` ~ `Light_Level`*Treatment, data = Pbur)

#Summary of the analysis
summary(res.aov_Colp_GR)
summary(res.aov_Pbur_GR)

TukeyHSD(res.aov_Pbur_GR)
```

##Tukey HSD for growth rate of Colpidium
```{r}
TukeyHSD(res.aov_Colp_GR)
```

##Tukey HSD for max population of P. bursaria
```{r}
# Compute the analysis of variance
res.aov_Colp_MP <- aov(`Cells/mL` ~ `Light_Level`*Treatment, data = Colp)
res.aov_Pbur_MP <- aov(`Cells/mL` ~ `Light_Level`*Treatment, data = Pbur)

# Summary of the analysis
summary(res.aov_Colp_MP)
summary(res.aov_Pbur_MP)

TukeyHSD(res.aov_Pbur_MP)
```

##Tukey HSD for max population of Colpidium
```{r}
TukeyHSD(res.aov_Colp_MP)
```


##Graphing Tukey HSD tests
Create the graphs to represent Tukey HSD test. The different letters represent statistical difference at the p > 0.05 level. 

```{r}
Pbur$Treatment = factor(Pbur$Treatment, labels = c("P. bursaria competition", "P. bursaria monoculture"))
Pbur= mutate(Pbur, Treatment=fct_relevel(Treatment,"P. bursaria monoculture","P. bursaria competition"))
```

```{r}
o=4 #labels font size
t=10 #title size
l=6 #legend text size
z=8 #tag size
legend=c(0.8, 0.7) #legend position
title=c(10,-10,-10,10) #title position for panels
lightcolorsP=c("orange4", "orange3","orange2", "goldenrod1") #colors for the graphs, light levels, light level gradient 
lightcolorsC=c("steelblue4","steelblue","lightsteelblue3", "slategray1") #colors for the graphs, light levels, light level gradient
# grayscale = c("#666666", "#999999","#CCCCCC", "#FFFFFF") use this for grayscale 
exp_days_color=c("steelblue4","lightsteelblue3","orange4","goldenrod1")
exp_days_linetype=c("solid", "longdash","solid", "longdash")
exp_days_shape=c(16,16,25,25)


##Labels for the bar graph, different values demonstrate statistical significance at the level of p > 0.05
p2<-c("a", "a", "b", "a", "b", "a", "a", "a") #bar graph labels max population for P. bur 
c2<-c("abc", "bc", "bd", "e", "c", "f", "d", "c")#bar graph labels max population Colp
pg2<- c(NA, NA, "b", "b", "b", "b", "b", "b")#bar graph labels growth rate for P. bur
pg2.1 <- c("a","c", NA,NA,NA,NA,NA,NA)
cg2<- c("abc", "cd", "bd", "ce", "c", "e", "d", "bcd")#bar graph labels growth rate Colp

```

A) Creating plot for P. bursaria graph maximum population. 
```{r}
Pbur_MP_ana<- summarySE(Pbur, measurevar="Cells/mL", groupvars=c("Light_Level", "Treatment")) #create analysis dataframe
Pbur_MP_ana$Light_Level = factor(Pbur_MP_ana$Light_Level, labels = c("0", "100", "200", "50"))
Pbur_MP_ana= mutate(Pbur_MP_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))

p.Pbur_MP <- ggplot(Pbur_MP_ana, aes(x = `Treatment`, y = `Cells/mL`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+   
    scale_fill_manual(values=lightcolorsP)+
    geom_text(aes(label = p2, y = `Cells/mL`+se), vjust = -.3, position = position_dodge(0.9),size=o, show.legend=FALSE) +
  # geom_text(aes(label=p2), vjust = -0.5, position = position_dodge(0.9), size=o)+
  geom_errorbar(aes(ymin=`Cells/mL`-se, ymax=`Cells/mL`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9),
                show.legend=FALSE) + #add axis and title names
    # scale_x_discrete("Treatment", labels=expression(paste(italic("P. bursaria")~"monoculture", italic("P. bursaria")~"competition")))+
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=8),
    axis.text.x=element_text(size=9,color='black',hjust=0.5),
    axis.title.x=element_blank(),
   axis.title.y=element_blank(),
   legend.position=("none"),
    legend.key.size = unit(0.7, 'lines'),
    legend.background=element_rect(colour=NA, fill=NA),
    # legend.position = c(0.85,0.85),
    legend.margin=margin(c(-5,1,1,1)),
    legend.title=element_text(size=9),
    legend.text=element_text(size=8),
   plot.title=element_text(size=t, margin=margin(title)))+
    labs(
       y="Max population (cells/mL)",
       fill= expression("Light intensity \n(μmol quanta m"^-2*"s"^-1*")"))

```

B) Creating plot for Colpidium maximum population. 
```{r}
Colp$Treatment = factor(Colp$Treatment, labels = c("Colpidium monoculture", "Colpidium competition"))
# Colp= mutate(Colp, Treatment=fct_relevel(Treatment,"Colpidium monoculture","Colpidium competition"))

Colp_MP_ana<- summarySE(Colp, measurevar="Cells/mL", groupvars=c("Light_Level", "Treatment")) #create analysis dataframe
Colp_MP_ana$Light_Level = factor(Colp_MP_ana$Light_Level, labels = c("0", "100", "200", "50"))
Colp_MP_ana= mutate(Colp_MP_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))

p.Colp_MP <- ggplot(Colp_MP_ana, aes(x = `Treatment`, y = `Cells/mL`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+  
  scale_fill_manual(values=lightcolorsC)+
  geom_text(aes(label = c2, y = `Cells/mL`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
  geom_errorbar(aes(ymin=`Cells/mL`-se, ymax=`Cells/mL`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
    # scale_x_discrete("Treatment", labels=expression(paste(italic("Colpidium")~"monoculture", italic("Colpidium")~"competition")))+
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=9, color='black', hjust = 0.5),
    axis.title.x=element_blank(),
   axis.title.y=element_text(size=9),
   legend.position=("none"),
    legend.key.size = unit(0.7, 'lines'),
    legend.background=element_rect(colour=NA, fill=NA),
    # legend.position = c(0.85,0.85),
    legend.margin=margin(c(-5,1,1,1)),
    legend.title=element_text(size=9),
    legend.text=element_text(size=8),
   plot.title=element_text(siz=t, margin = margin(title)),
    plot.tag = element_text(size=z))+
    labs(
        y="Max population (cells/mL)",
         fill= expression("Light intensity \n(μmol quanta m"^-2*"s"^-1*")"))



```


```{r}
o=4 #labels font size
t=10 #title size
z=8 #tag size
legend=c(0.8, 0.7) #legend position

##The following labels are based on the Tukey's HSD test from above, where different letters represent statistically significant differences at the p < 0.05 level

p<-c("a", "a", "a", "a", "b", "a", "a", "a", "b", "a", "a", "a", "a", "a", "a", "a")#bar graph labels for max pop P bur
c<-c("ab", "bc", "c", "ab", "a", "ab", "bc", "a", "ab", "c", "ab", "ab", "c", "c")#bar graph labels for max pop Colp
pg<- c("a", "d", "d", "a", "b", "b", "ce", "bc", "b", "b", "ce", "bc", "bc", "b", "e", "b")#bar graph labels for growth rate P bur
cg<- c("abd", "ab", "ad", "abd", "b", "ab", "d", "c", "abd", "d", "abd", "abd", "ab", "abd")#bar graph labels for growth rate Colp
```

C) Creating plot for P. bursaria growth rate. 
```{r}
##First create summary SE for growth rate (above was summary SE for max populations)
Pbur_GR_ana<- summarySE(Pbur, measurevar="Growth_Rate", groupvars=c("Light_Level", "Treatment")) #create analysis df 
Pbur_GR_ana$Light_Level = factor(Pbur_GR_ana$Light_Level, labels = c("0μE", "100μE", "200μE", "50μE"))
Pbur_GR_ana= mutate(Pbur_GR_ana, Light_Level=fct_relevel(Light_Level,"0μE","50μE", "100μE", "200μE"))

p.Pbur_GR <- ggplot(Pbur_GR_ana, aes(x = `Treatment`, y = `Growth_Rate`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+
  geom_text(aes(label = pg2, y = `Growth_Rate`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
  geom_text(aes(label = pg2.1, y = 0), vjust = -.3 , position = position_dodge(0.9),size=o) +
  scale_fill_manual(values=lightcolorsP)+
  geom_errorbar(aes(ymin=`Growth_Rate`-se, ymax=`Growth_Rate`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
    # scale_x_discrete("Treatment", labels=expression(paste(italic("P. bursaria")~"monoculture", italic("P. bursaria")~"competition")))+
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=9),
    axis.text.x=element_text(size=9, color='black', hjust=0.5),
    # axis.text.x=element_blank(),
    axis.title.x=element_blank(),
   axis.title.y=element_blank(),
   plot.title=element_text(size=t, margin = margin(title)),
    legend.position=("none"))+
       labs(
            y="Growth rate (per day)")
```

D) Create plot for Colpidium growth rate. 
```{r}
Colp_GR_ana<- summarySE(Colp, measurevar="Growth_Rate", groupvars=c("Light_Level", "Treatment")) #create analysis df 
Colp_GR_ana$Light_Level = factor(Colp_GR_ana$Light_Level, labels = c("0μE", "100μE", "200μE", "50μE"))
Colp_GR_ana= mutate(Colp_GR_ana, Light_Level=fct_relevel(Light_Level,"0μE","50μE", "100μE", "200μE"))

p.Colp_GR <- ggplot(Colp_GR_ana, aes(x = `Treatment`, y = `Growth_Rate`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+  
  scale_fill_manual(values=lightcolorsC)+
  geom_errorbar(aes(ymin=`Growth_Rate`-se, ymax=`Growth_Rate`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
  geom_text(aes(label = cg2, y = `Growth_Rate`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
 
  theme_light()+
  theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=9),
    axis.text.x=element_text(size=9, color='black'),
    axis.title.y=element_text(size=9), 
    axis.title=element_blank(),
    legend.text=element_text(size=l),
    plot.title=element_text(size=t, margin=margin(title)),
    legend.position=("none"))+
    labs(
            y="Growth rate (per day)")
```


##Graphs for population dynamics of P. bursaria and Colpidium
```{r, warning=FALSE}
library(cowplot)
library(ggpubr)

p.Pbur_MP <- p.Pbur_MP + ylim(c(0, 7000)) 
p.Pbur_GR <- p.Pbur_GR + ylim(c(-0.22, 0.2))
p.Colp_MP <- p.Colp_MP + ylim(c(0, 3000))
p.Colp_GR <- p.Colp_GR + ylim(c(0, 0.9))


light<- ggdraw() +
  draw_plot(p.Colp_MP, 0, .5, .5, .49) +
  draw_plot(p.Colp_GR, 0, 0, .5, .49) +
  draw_plot(p.Pbur_MP, .5, .5, .5, .49) +
  draw_plot(p.Pbur_GR, .5, 0, .5, .49)
  # draw_plot_label(label = c("P. bursaria", "Colpidium"),  fontfamily="sans", fontface="plain",size = 10, x=c(-.03,.47), y=c(1,1))
light
# pdf("plot.pdf", width=8, height=5.5)
# plot(light)

```

#Appendix S1: Figure S2
We analyzed the experimental day at which each treatment (Colpidium monoculture, Colpidium competition, P. bursaria monoculture, and P. bursaria competition) reached maximum population at each light level (0, 50, 100 and 200 mol quanta m^-2 s^-1). 

```{r}
Pbur_ED_ana<- summarySE(Pbur, measurevar="Experimental_Day", groupvars=c("Light_Level", "Species",'Treatment'))
Pbur_ED_ana$Light_Level = factor(Pbur_ED_ana$Light_Level, labels = c("0", "100", "200", "50"))
Pbur_ED_ana= mutate(Pbur_ED_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))


Colp_ED_ana<- summarySE(Colp, measurevar="Experimental_Day", groupvars=c("Light_Level", "Species",'Treatment'))
Colp_ED_ana$Light_Level = factor(Colp_ED_ana$Light_Level, labels = c("0", "100", "200", "50"))
Colp_ED_ana= mutate(Colp_ED_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))

name <- c(expression(paste(italic("Colpidium")~"monoculture")),
         expression(paste(italic("Colpidium")~"competition")),
          expression(paste(italic("P. bursaria")~"monoculture")),
         expression(paste(italic("P. bursaria")~"competition")))

ED_ana<- rbind(Colp_ED_ana,Pbur_ED_ana)

maxpop <- ggplot(ED_ana, aes (x = `Light_Level`, y = `Experimental_Day`, color=`Treatment`, fill=`Treatment`, linetype=`Treatment`)) +
  geom_point(binaxis='y', stackdir='center', size=2.5, aes(color=`Treatment`, fill=`Treatment`, shape=`Treatment`))+
  geom_errorbar(aes(ymin=`Experimental_Day`-se, ymax=`Experimental_Day`+se, color=`Treatment`),
                  width=(.1))+               # Width of the error bars
                  # position=position_dodge(.5))+  #add axis and title names
  geom_line(aes(group=`Treatment`, color=`Treatment`, linetype=`Treatment`))+
  scale_color_manual(values=exp_days_color,
                     labels=name)+
  scale_linetype_manual(values=exp_days_linetype,
                        labels=name)+
  scale_shape_manual(values=exp_days_shape,
                     labels=name)+
  scale_fill_manual(values=exp_days_color,
                    labels=name)+
  theme_light()+
  theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=12),
    axis.text.x=element_text(size=12, color='black'),
    # axis.title.y=element_text(size=12), 
    # axis.title.x=element_text(size=12),
    axis.title=element_text(size=12),
    legend.text=element_text(size=12),
    legend.key.width=unit(1,"cm"),
    legend.text.align = 0,
    legend.title=element_blank(),
    plot.title=element_text(size=t, margin=margin(title)))+
    labs(y="Experimental Day",
       x= expression(paste("Light intensity (",italic("μ"),"mol quanta m"^-2*"s"^-1*")")))

maxpop<- ggdraw() +
  draw_plot(maxpop, 0, .5, 0.8, .5)

maxpop
# pdf("maxpop.pdf", width=8, height=5)
# plot(maxpop)
```
