---
title: Niche expansion via acquired metabolism facilitates competitive dominance in
  planktonic communities
author: "Veronica Hsu, F. Pfab, and Holly V. Moeller"
date: "7/9/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###
We performed a Tukey HSD test to determine statistically significance differences at the p < 0.05 level, with independent variables being light (at 0, 50, 100, 200 μmol quanta m-2 s-1) and treatment (monoculture vs competition) and the dependent variables were max populations and growth rates. 

#Tukey HSD for growth rate and maximum population for Colpidium and P. bursaria.
Below is how we generated the datafile "AcqMet_Tukey_v1" using res.aov function and experimental results of P. bursaria and Colpidium.

##Tukey HSD for growth rate of P. bursaria 
```{r}
# # Compute the analysis of variance
# res.aov_Colp_GR <- aov(`Growth_Rate` ~ `Light_Level`*Treatment, data = Colp) #data = Colp, "Colp"" is a datafile with all data on Colpidium population dynamics from datafile Pbur_Colp_v1.csv
# res.aov_Pbur_GR<- aov(`Growth_Rate` ~ `Light_Level`*Treatment, data = Pbur) #data = Pbur, "Pbur" is a datafile with all data on Colpidium population dynamics
# 
# #Summary of the analysis
# summary(res.aov_Colp_GR)
# summary(res.aov_Pbur_GR)
# 
# TukeyHSD(res.aov_Pbur_GR)
```

##Tukey HSD for growth rate of Colpidium
```{r}
# TukeyHSD(res.aov_Colp_GR)
```

#Tukey HSD for max population of P. bursaria
```{r}
# # Compute the analysis of variance
# res.aov_Colp_MP <- aov(`Cells` ~ `Light_Level`*Treatment, data = Colp)
# res.aov_Pbur_MP <- aov(`Cells` ~ `Light_Level`*Treatment, data = Pbur)
# 
# # Summary of the analysis
# summary(res.aov_Colp_MP)
# summary(res.aov_Pbur_MP)
# 
# TukeyHSD(res.aov_Pbur_MP)
```

#Tukey HSD for max population of Colpidium
```{r}
# TukeyHSD(res.aov_Colp_MP)
```


Preparation: Load the data
```{r}
library(readr)
library(readr)
library(forcats)
library(ggplot2)
library(Rmisc)
library(dplyr)
library(tidyr)
df <- read_csv("AcqMet_Tukey_v1")

df$`Light_Level`<- as.character(df$`Light_Level`)
df$Treatment <- as.character(df$Treatment)
df$Cells <- as.double(df$Cells)
df$Growth_Rate <- as.double(df$Growth_Rate)

Pbur <- df[df$Species == "P bur", ]
Colp <- df[df$Species == "Colp", ]


Pbur$Treatment = factor(Pbur$Treatment, labels = c("P. bursaria competition","P. bursaria monoculture"))
Colp$Treatment = factor(Colp$Treatment, labels = c("Colpidium competition","Colpidium monoculture")) 
Pbur= mutate(Pbur, Treatment=fct_relevel(Treatment,"P. bursaria monoculture","P. bursaria competition"))
Colp= mutate(Colp, Treatment=fct_relevel(Treatment,"Colpidium monoculture","Colpidium competition"))
```


##Graphing Tukey HSD tests
Create the graphs to represent Tukey HSD test. The different letters represent statistical difference at the p > 0.05 level. 
```{r}
o=4 #labels font size
t=10 #title size
l=6 #legend text size
z=8 #tag size
legend=c(0.8, 0.7) #legend position
title=c(10,-10,-10,10) #title position for panels
lightcolorsP=c("orange4", "orange3","orange2", "goldenrod1") #colors for the graphs, light levels, light level gradient 
lightcolorsC=c("steelblue4","steelblue","lightsteelblue3", "slategray1") #colors for the graphs, light levels, light level gradient
# grayscale = c("#666666", "#999999","#CCCCCC", "#FFFFFF") use this for grayscale 


##Labels for the bar graph, different values demonstrate statistical significance at the level of p > 0.05
p2<-c("a", "a", "b", "a", "b", "a", "a", "a") #bar graph labels max population for P. bur 
c2<-c("abc", "bc", "bd", "e", "c", "f", "d", "c")#bar graph labels max population Colp
pg2<- c(NA, NA, "b", "b", "b", "b", "b", "b")#bar graph labels growth rate for P. bur
pg2.1 <- c("a","c", NA,NA,NA,NA,NA,NA)
cg2<- c("abc", "cd", "bd", "ce", "c", "e", "d", "bcd")#bar graph labels growth rate Colp

```

A) Creating plot for P. bursaria graph maximum population. 
```{r}
Pbur_MP_ana<- summarySE(Pbur, measurevar="Cells", groupvars=c("Light_Level", "Treatment")) #create analysis dataframe
Pbur_MP_ana$Light_Level = factor(Pbur_MP_ana$Light_Level, labels = c("0", "100", "200", "50"))
Pbur_MP_ana= mutate(Pbur_MP_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))

p.Pbur_MP <- ggplot(Pbur_MP_ana, aes(x = `Treatment`, y = `Cells`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+   
    scale_fill_manual(values=lightcolorsP)+
    geom_text(aes(label = p2, y = `Cells`+se), vjust = -.3, position = position_dodge(0.9),size=o, show.legend=FALSE) +
  # geom_text(aes(label=p2), vjust = -0.5, position = position_dodge(0.9), size=o)+
  geom_errorbar(aes(ymin=`Cells`-se, ymax=`Cells`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9),
                show.legend=FALSE) + #add axis and title names
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=8),
    axis.text.x=element_text(size=8,color='black'),
    # axis.text.x=element_blank(),
    axis.title.x=element_blank(),
   axis.title.y=element_text(size=8),
   legend.position=("none"),
    legend.key.size = unit(0.7, 'lines'),
    legend.background=element_rect(colour=NA, fill=NA),
    # legend.position = c(0.85,0.85),
    legend.margin=margin(c(-5,1,1,1)),
    legend.title=element_text(size=8),
    legend.text=element_text(size=8),
   plot.title=element_text(size=t, margin=margin(title)))+
    labs(title="  A",
       y="Max population (cells/mL)",
       fill= expression("Light intensity \n(μmol quanta m"^-2*"s"^-1*")"))

```

B) Creating plot for Colpidium maximum population. 
```{r}
Colp_MP_ana<- summarySE(Colp, measurevar="Cells", groupvars=c("Light_Level", "Treatment")) #create analysis dataframe
Colp_MP_ana$Light_Level = factor(Colp_MP_ana$Light_Level, labels = c("0", "100", "200", "50"))
Colp_MP_ana= mutate(Colp_MP_ana, Light_Level=fct_relevel(Light_Level,"0","50", "100", "200"))

p.Colp_MP <- ggplot(Colp_MP_ana, aes(x = `Treatment`, y = `Cells`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+  
  scale_fill_manual(values=lightcolorsC)+
  geom_text(aes(label = c2, y = `Cells`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
  geom_errorbar(aes(ymin=`Cells`-se, ymax=`Cells`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=8),
    axis.text.x=element_text(size=8, color='black'),
    axis.title.x=element_blank(),
   axis.title.y=element_blank(),
   legend.position=("none"),
    legend.key.size = unit(0.7, 'lines'),
    legend.background=element_rect(colour=NA, fill=NA),
    # legend.position = c(0.85,0.85),
    legend.margin=margin(c(-5,1,1,1)),
    legend.title=element_text(size=8),
    legend.text=element_text(size=8),
   plot.title=element_text(siz=t, margin = margin(title)),
    plot.tag = element_text(size=z))+
    labs(title = "  B",
         fill= expression("Light intensity \n(μmol quanta m"^-2*"s"^-1*")"))



```

```{r}
o=4 #labels font size
t=10 #title size
z=8 #tag size
legend=c(0.8, 0.7) #legend position

##The following labels are based on the Tukey's HSD test from above, where different letters represent statistically significant differences at the p < 0.05 level

p<-c("a", "a", "a", "a", "b", "a", "a", "a", "b", "a", "a", "a", "a", "a", "a", "a")#bar graph labels for max pop P bur
c<-c("ab", "bc", "c", "ab", "a", "ab", "bc", "a", "ab", "c", "ab", "ab", "c", "c")#bar graph labels for max pop Colp
pg<- c("a", "d", "d", "a", "b", "b", "ce", "bc", "b", "b", "ce", "bc", "bc", "b", "e", "b")#bar graph labels for growth rate P bur
cg<- c("abd", "ab", "ad", "abd", "b", "ab", "d", "c", "abd", "d", "abd", "abd", "ab", "abd")#bar graph labels for growth rate Colp
```

C) Creating plot for P. bursaria growth rate. 
```{r}
##First create summary SE for growth rate (above was summary SE for max populations)
Pbur_GR_ana<- summarySE(Pbur, measurevar="Growth_Rate", groupvars=c("Light_Level", "Treatment")) #create analysis df 
Pbur_GR_ana$Light_Level = factor(Pbur_GR_ana$Light_Level, labels = c("0μE", "100μE", "200μE", "50μE"))
Pbur_GR_ana= mutate(Pbur_GR_ana, Light_Level=fct_relevel(Light_Level,"0μE","50μE", "100μE", "200μE"))

p.Pbur_GR <- ggplot(Pbur_GR_ana, aes(x = `Treatment`, y = `Growth_Rate`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+
  geom_text(aes(label = pg2, y = `Growth_Rate`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
  geom_text(aes(label = pg2.1, y = 0), vjust = -.3 , position = position_dodge(0.9),size=o) +
  scale_fill_manual(values=lightcolorsP)+
  geom_errorbar(aes(ymin=`Growth_Rate`-se, ymax=`Growth_Rate`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
  theme_light()+
   theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=8),
    # axis.text.x=element_text(size=8, color='black'),
    axis.text.x=element_text(size=8, color='black'),
    axis.title.x=element_blank(),
   axis.title.y=element_text(size=8),
   plot.title=element_text(size=t, margin = margin(title)),
    legend.position=("none"))+
       labs(title = "  C",
            y="Growth rate (per day)")
```

D) Create plot for Colpidium growth rate. 
```{r}
Colp_GR_ana<- summarySE(Colp, measurevar="Growth_Rate", groupvars=c("Light_Level", "Treatment")) #create analysis df 
Colp_GR_ana$Light_Level = factor(Colp_GR_ana$Light_Level, labels = c("0μE", "100μE", "200μE", "50μE"))
Colp_GR_ana= mutate(Colp_GR_ana, Light_Level=fct_relevel(Light_Level,"0μE","50μE", "100μE", "200μE"))

p.Colp_GR <- ggplot(Colp_GR_ana, aes(x = `Treatment`, y = `Growth_Rate`, fill = `Light_Level`)) +
  geom_col(position = "dodge")+  
  scale_fill_manual(values=lightcolorsC)+
  geom_errorbar(aes(ymin=`Growth_Rate`-se, ymax=`Growth_Rate`+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) + #add axis and title names
  geom_text(aes(label = cg2, y = `Growth_Rate`+se), vjust = -.3, position = position_dodge(0.9),size=o) +
  theme_light()+
  theme(axis.line = element_line(colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y=element_text(size=8),
    # axis.text.x=element_blank(),
    axis.text.x=element_text(size=8, color='black'),
    axis.title.y=element_blank(), #x axis blank
    axis.title=element_blank(),
    legend.text=element_text(size=l),
    plot.title=element_text(size=t, margin=margin(title)),
    legend.position=("none"))+
    labs(title="  D")
```


##Graphs for population dynamics of P. bursaria and Colpidium
```{r, warning=FALSE}
library(cowplot)
library(ggpubr)

p.Pbur_MP <- p.Pbur_MP + ylim(c(0, 7000)) 
p.Pbur_GR <- p.Pbur_GR + ylim(c(-0.22, 0.2))
p.Colp_MP <- p.Colp_MP + ylim(c(0, 3000))
p.Colp_GR <- p.Colp_GR + ylim(c(0, 0.9))


light<- ggdraw() +
  draw_plot(p.Pbur_MP, 0, .5, .5, .49) +
  draw_plot(p.Pbur_GR, 0, 0, .5, .49) +
  draw_plot(p.Colp_MP, .5, .5, .5, .49) +
  draw_plot(p.Colp_GR, .5, 0, .5, .49)
  # draw_plot_label(label = c("P. bursaria", "Colpidium"),  fontfamily="sans", fontface="plain",size = 10, x=c(-.03,.47), y=c(1,1))
light
pdf("plot.pdf", width=8, height=5)
plot(light)


```


